msqrob2TMT: robust linear mixed models for inferring differential abundant proteins in labelled experiments with arbitrarily complex design

### 2025-02-05 paper/manuscript.qmd
Changed figure names

knitr::include_graphics("../figs/model_overview.png", error = FALSE)

@@

```{r, echo  =FALSE, message  =FALSE, warning = FALSE, eval=TRUE, out.width="100%"}
#| label: fig-WorkflowComparisons
#| fig-cap: "Performance in the MSstatsTMT spike-in study. Panel A: True positive rate (TPR) - false discovery proportion (FDP) plots for DEqMS, MSstatsTMT, msTrawler and msqrob2TMT workflows. Note, that the import function of msTrawler did not properly import the data. So we also provide results for msTrawler with a refactored import function, referred to as msTrawler\_fixed. The combined performance over all the dilution comparisons is shown. Dots indicate the TPR and FDP obtained at the 5% FDR threshold. Panel B: Number of fitted proteins for each workflow. Panel C: Number of true and false positives over all the comparisons."
#| fig-pos: 'H'
knitr::include_graphics("../figs/spikein1_tpr_fdp.png", error = FALSE)

@@

```{r, echo  =FALSE, warning = FALSE, message  =FALSE, out.width="100%"}
#| label: fig-fold_change_spike
#| fig-cap: "Boxplots showing the log2 FC distributions for the spike-in proteins, focusing on two comparisons with the highest and lowest difference in spike-in concentrations. Fold changes are estimated by DEqMS, msqrob2TMT, MSstatsTMT, and msTrawler workflows. The grey dotted line is the true log2 FC for the comparison. Panel A: non spike-in proteins (yeast); Panel B: spike-in proteins."
#| fig-pos: 'H'
knitr::include_graphics("../figs/spikein1_logFC.png", error = FALSE)
```

@@

knitr::include_graphics("../figs/spikein1_msqrob2_estimation_features.png", error = FALSE)

@@

knitr::include_graphics("../figs/spikein2_tpr_fdp.png", error = FALSE)
```

@@

```{r, echo  =FALSE, warning = FALSE, message  =FALSE, out.width="100%"}
#| label: fig-fold_change_spike2
#| fig-cap: "Boxplots showing the log2 FC distributions for the spike-in proteins, focusing on two comparisons with the highest and lowest difference in spike-in concentrations. Fold changes are estimated by DEqMS, msqrob2TMT, and msTrawler workflows. The grey dotted line is the true log2 FC for the comparison. Panel A: non-spiked proteins (mouse); Panel B: spike-in proteins (yeast)."
#| fig-pos: 'H'
knitr::include_graphics("../figs/spikein2_logFC.png", error = FALSE)
```

@@

\section*{Spike-In Dataset 1}
```{r, echo  =FALSE, message  =FALSE, warning = FALSE, eval=TRUE,out.width="90%"}
#| label: fig-WorkflowComparisonsFull
#| fig-cap: "True positive rate (TPR) - false discovery proportion (FDP) plots for DEqMS, msqrob2TMT, msTrawler, msTrawler\_fixed with a refactored import function and MSstatsTMT workflows based on all pairwise spike-in comparisons. Full Caption on the next page."
#| fig-pos: 'H'
knitr::include_graphics("../figs/spikein1_supp_tpr_fdp.png", error = FALSE)
```

@@

```{r, echo  =FALSE, message  =FALSE, warning = FALSE, eval=TRUE, out.width="100%"}
#| label: fig-WorkflowComparisonsSeparateA
#| fig-cap: "True positive rate (TPR) - false discovery proportion (FDP) plots for DEqMS, MSstatsTMT, msTrawler, msTrawler\_fixed with a refactored import function and  msqrob2TMT workflows. The performance is based on the results that are returned by each workflow. Dots indicate the TPR and FDP obtained at the 5% FDR threshold."
#| fig-pos: 'H'
knitr::include_graphics("../figs/spikein1_supp_comparison_1.png", error = FALSE)
```

```{r, echo  =FALSE, message  =FALSE, warning = FALSE, eval=TRUE, out.width="100%"}
#| label: fig-WorkflowComparisonsSeparateB
#| fig-cap: "True positive rate (TPR) - false discovery proportion (FDP) plots for DEqMS, MSstatsTMT, msTrawler, msTrawler\_fixed with a refactored import function and  msqrob2TMT workflows. The TPR is based on all ground truth DA proteins as the maximum number of true positives that can be reported for each comparison (40 spike-in UPS proteins). Dots indicate the TPR and FDP obtained at the 5% FDR threshold."
#| fig-pos: 'H'
knitr::include_graphics("../figs/spikein1_supp_comparison_2.png", error = FALSE)
```

@@

knitr::include_graphics("../figs/spikein1_supp_comparison_3.png", error = FALSE)
```

@@

knitr::include_graphics("../figs/spikein1_supp_tpr_fdp_replication.png", error = FALSE)
```

```{r, echo  =FALSE, message  =FALSE, warning = FALSE, out.width="100%"}
#| label: fig-spikein1_pvalue_dist_nondiff
#| fig-cap: "Histograms of the p-values from non-spike-in proteins for the DEqMS, msqrob2TMT, MSstatsTMT and msTrawler workflows for two distinct pairwise comparisons between spike-in dilutions."
#| fig-pos: 'H'
knitr::include_graphics("../figs/spikein1_null_distribution.png", error = FALSE)
```

@@

knitr::include_graphics("../figs/spikein1_supp_compare_preprocess.png", error = FALSE)
```

@@

knitr::include_graphics("../figs/spikein2_supp_tpr_fdp.png", error = FALSE)
```

@@

knitr::include_graphics("../figs/spikein2_supp_compare_preprocess.png", error = FALSE)
```

```{r, echo  =FALSE, message  =FALSE, warning = FALSE, out.width="100%"}
#| label: fig-spikein2_pvalue_dist_nondiff
#| fig-cap: "Histograms of the p-values from non-spike-in proteins for the DEqMS, msqrob2TMT, and msTrawler workflows for four distinct pairwise comparisons between spike-in dilutions."
#| fig-pos: 'H'
knitr::include_graphics("../figs/spikein2_null_distribution.png", error = FALSE)
```

@@

knitr::include_graphics("../figs/mouse_upset_early.png", error = FALSE)

@@

knitr::include_graphics("../figs/mouse_upset_late.png", error = FALSE)

@@

knitr::include_graphics("../figs/mouse_upset_average.png", error = FALSE)
```

```{r, eval=FALSE, echo  =FALSE, message  =FALSE, warning = FALSE, out.width="100%"}
#| label: fig-Upset-int
#| fig-cap: "Upset plot showing the overlap in significant proteins between the MSstatsTMT and the msqrob2TMT workflows for the interaction."
#| fig-pos: 'H'

knitr::include_graphics("../figs/mouse_upset_interaction.png", error = FALSE)

@@

knitr::include_graphics("../figs/mouse_supp_ora.png", error = FALSE)

```{r, echo  =FALSE, message  =FALSE, warning = FALSE,out.width="100%"}
#| label: fig-mouse-mock-pval
#| fig-cap: Histogram of the p-values from the mock analysis on the mouse dataset.  Only results for our PSM-level and protein-level msqrob2TMT workflows are shown, as the state-of-the-art tools could not fit the appropriate models to the data of the mouse study after including an additional mock treatment."
knitr::include_graphics("../figs/mouse_supp_null_distribution.png", error = FALSE)
```

>>>>

knitr::include_graphics("../figs/figure1.png", error = FALSE)

@@

```{r, echo  =FALSE, message  =FALSE, warning = FALSE, eval=TRUE, out.width="100%"}
#| label: fig-WorkflowComparisons
#| fig-cap: "Performance in the MSstatsTMT spike-in study. Panel A: True positive rate (TPR) - false discovery proportion (FDP) plots for DEqMS, MSstatsTMT, msTrawler and msqrob2TMT workflows. Note, that the import function of msTrawler did not properly import the data. So we also provide results for msTrawler with a refactored import function, referred to as msTrawler\_fixed. The combined performance over all the dilution comparisons is shown. Dots indicate the TPR and FDP obtained at the 5% FDR threshold. Panel B: Number of fitted proteins for each workflow. Panel C: Number of true and false positives over all the comparisons."
#| fig-pos: 'H'
knitr::include_graphics("../figs/figure2.png", error = FALSE)

@@

```{r, echo  =FALSE, warning = FALSE, message  =FALSE, out.width="100%"}
#| label: fig-fold_change_spike
#| fig-cap: "Boxplots showing the log2 FC distributions for the spike-in proteins, focusing on two comparisons with the highest and lowest difference in spike-in concentrations. Fold changes are estimated by DEqMS, msqrob2TMT, MSstatsTMT, and msTrawler workflows. The grey dotted line is the true log2 FC for the comparison. Panel A: non spike-in proteins (yeast); Panel B: spike-in proteins."
#| fig-pos: 'H'
knitr::include_graphics("../figs/figure3.png", error = FALSE)
```

@@

knitr::include_graphics("../figs/figure4.png", error = FALSE)

@@

knitr::include_graphics("../figs/figure5.png", error = FALSE)
```

@@

```{r, echo  =FALSE, warning = FALSE, message  =FALSE, out.width="100%"}
#| label: fig-fold_change_spike2
#| fig-cap: "Boxplots showing the log2 FC distributions for the spike-in proteins, focusing on two comparisons with the highest and lowest difference in spike-in concentrations. Fold changes are estimated by DEqMS, msqrob2TMT, and msTrawler workflows. The grey dotted line is the true log2 FC for the comparison. Panel A: non-spiked proteins (mouse); Panel B: spike-in proteins (yeast)."
#| fig-pos: 'H'
knitr::include_graphics("../figs/figure6.png", error = FALSE)
```

@@

\section*{Spike-In Dataset 1}
```{r, echo  =FALSE, message  =FALSE, warning = FALSE, eval=TRUE,out.width="90%"}
#| label: fig-WorkflowComparisonsFull
#| fig-cap: "True positive rate (TPR) - false discovery proportion (FDP) plots for DEqMS, msqrob2TMT, msTrawler, msTrawler\_fixed with a refactored import function and MSstatsTMT workflows based on all pairwise spike-in comparisons. Full Caption on the next page."
#| fig-pos: 'H'
knitr::include_graphics("../figs/figureS1.png", error = FALSE)
```

@@

```{r, echo  =FALSE, message  =FALSE, warning = FALSE, eval=TRUE, out.width="100%"}
#| label: fig-WorkflowComparisonsSeparateA
#| fig-cap: "True positive rate (TPR) - false discovery proportion (FDP) plots for DEqMS, MSstatsTMT, msTrawler, msTrawler\_fixed with a refactored import function and  msqrob2TMT workflows. The performance is based on the results that are returned by each workflow. Dots indicate the TPR and FDP obtained at the 5% FDR threshold."
#| fig-pos: 'H'
knitr::include_graphics("../figs/figureS2.png", error = FALSE)
```

```{r, echo  =FALSE, message  =FALSE, warning = FALSE, eval=TRUE, out.width="100%"}
#| label: fig-WorkflowComparisonsSeparateB
#| fig-cap: "True positive rate (TPR) - false discovery proportion (FDP) plots for DEqMS, MSstatsTMT, msTrawler, msTrawler\_fixed with a refactored import function and  msqrob2TMT workflows. The TPR is based on all ground truth DA proteins as the maximum number of true positives that can be reported for each comparison (40 spike-in UPS proteins). Dots indicate the TPR and FDP obtained at the 5% FDR threshold."
#| fig-pos: 'H'
knitr::include_graphics("../figs/figureS3.png", error = FALSE)
```

@@

knitr::include_graphics("../figs/figureS4.png", error = FALSE)
```

@@

knitr::include_graphics("../figs/figureS5.png", error = FALSE)
```

```{r, echo  =FALSE, message  =FALSE, warning = FALSE, out.width="100%"}
#| label: fig-spikein1_pvalue_dist_nondiff
#| fig-cap: "Histograms of the p-values from non-spike-in proteins for the DEqMS, msqrob2TMT, MSstatsTMT and msTrawler workflows for two distinct pairwise comparisons between spike-in dilutions."
#| fig-pos: 'H'
knitr::include_graphics("../figs/figureS6.png", error = FALSE)
```

@@

knitr::include_graphics("../figs/figureS7.png", error = FALSE)
```

@@

knitr::include_graphics("../figs/figureS8.png", error = FALSE)
```

@@

knitr::include_graphics("../figs/figureS9.png", error = FALSE)
```

```{r, echo  =FALSE, message  =FALSE, warning = FALSE, out.width="100%"}
#| label: fig-spikein2_pvalue_dist_nondiff
#| fig-cap: "Histograms of the p-values from non-spike-in proteins for the DEqMS, msqrob2TMT, and msTrawler workflows for four distinct pairwise comparisons between spike-in dilutions."
#| fig-pos: 'H'
knitr::include_graphics("../figs/figureS10.png", error = FALSE)
```

@@

knitr::include_graphics("../figs/figureS11.png", error = FALSE)

@@

knitr::include_graphics("../figs/figureS12.png", error = FALSE)

@@

knitr::include_graphics("../figs/figureS13.png", error = FALSE)
```

knitr::include_graphics("../figs/figureS14.png", error = FALSE)

@@

```{r, echo  =FALSE, message  =FALSE, warning = FALSE,out.width="100%"}
#| label: fig-mouse-mock-pval
#| fig-cap: Histogram of the p-values from the mock analysis on the mouse dataset.  Only results for our PSM-level and protein-level msqrob2TMT workflows are shown, as the state-of-the-art tools could not fit the appropriate models to the data of the mouse study after including an additional mock treatment."
knitr::include_graphics("../figs/figureS15.png", error = FALSE)
```
